Large scale agile transformations: A literature review

### 2013-01-25 tex/article.tex
Optimizing, removing text

It is typical for large companies to function according to plan-driven software
engineering models [Cite?!] (also why?). These models strive for consistent quality and performance
by rigorous planning and definition of process. This kind of approach is however
badly suited for software development, as projects typically encounter
situations that are too hard or impossible to foretell \citegen{Schwaber2002}.
The main problems in plan-driven development are the high cost of changes and
late feedback on quality \citegen{Petersen2010}. Long release cycles, cost of
resopnding to change and distance from customers undermine the  competitiveness
of companies. Agile methods are believed to bring remedy to these problems.

@@

Introducing agile methods in large organizations is more difficult than it is in
small organizations \citegen{Dyba2008}. The difficulty is partly related to size
bringing higher organizational inertia which slows down organizational change
\citegen{Livermore2008}. Agile methods are not founded on the use of individual
tools or practices, but rather on a holistic change in the way of thinking.
Adopting agile requires often change of the entire organizational culture
\citegen{Misra2010}.

One significant difference between small and large adoptions is that large
organizations have more dependencies between projects and teams. This increases
the need for formal documentation and thus reduces agility \citegen{Lindvall2004}.
In addition to inter-team coordination large organizations require development
teams to interact with other organizational units, which are often non-agile of
their nature. For instance, HR may demand individuals to have strictly specified
roles in projects \citegen{Boehm2005}, or a change control board may inhibit use of
continuous integration or refactoring \citegen{Lindvall2004}. All units affected by
the agile transformation need to be informed and heard, and the agile process
must be adjusted according to their needs \citegen{Lindvall2004, Cohn2003,
Boehm2005}.

Agile methods will also affect management and business related functions. A key
challenge is that management must move away from life cycle models and towards
iterative and feature centric models \citegen{Nerur2005}. The focus must be shifted
from a large scale scope to shorter term project planning \citegen{Misra2010}, as
agile methods emphasize that meaningful planning can only be done for the near
future \citegen{Cohn2003}. However, the lack of planning can be a concern as
business and customer relationships often build on long term roadmapping.
Enabling operation with shorter term planning requires educating stakeholders
and reviewing contracting practices \citegen{Boehm2005}.

@@

The literature review was conducted as an application of Kitchenham's
method for systematic literature review (SLR) \citegen{Kitchenham2007}. The method
was applied to be performed by only one researcher, and to serve as a
preliminary work evaluating the feasibility of a more in-depth review. The
required effort was limited by selecting only the following subset of SLR steps:
identification of primary sources, data extraction and synthesis. Steps left out
were the development of a systematic review protocol, strict study quality
assessment, and reviews in the study selection process.

@@

@@

\begin{table}[h]
    \begin{tabular}{ l@{ \hskip 0.4cm } l }
        \toprule
        Facet                  & Keywords   \\ \midrule
        Agile methods          & agile, scrum, lean, xp \\ 
        Organizational change  & transformation, transition, change, migration \\
        Large organization     & enterprise, organization, (large \texttt{AND} scale) \\
        \bottomrule
    \end{tabular}
    \caption{Facets and related search terms used}
    \label{table:searchterms}
\end{table}

The selection process of the primary studies is summarized in Figure
\ref{fig:selection_process}. The database search yielded 1603 matches from which
100 sources were chosen by the title. 17 unmatched sources from the preliminary
search were also included. Finally, 31 sources were selected based on the
abstract to be included for data extraction. One source was not available, which
reduced the final number of primary studies to 30.

\begin{figure}[htb]
  \begin{center}
    \includegraphics[width=0.5\textwidth]{researchprocess}
    \caption{Process of selecting primary studies}
    \label{fig:selection_process}
  \end{center}
\end{figure}

@@

The 30 primary sources were selected by four criteria: study type, size of
organization, use of agile methods in software development, and viewpoint on
transformation. Firstly, only case studies and experience reports on agile
software development were included in this study. Papers discussing agile
methods or organizational transformation on as a general topic were left out.
Secondly, the size of the organization described in the included primary studies
was to be large enough to show the challenges of large scale agile development
\citegen{Lindvall2004}. Further, only studies discussing agile software development
were included, leaving out for instance studies on agile management. Finally, to
be included each study had to provide a viewpoint on the organizational
transformation.

@@

The content of the resutls was guided by the data extraction form. The form
included some quantitative fields including \textit{start time of
transformation} and \textit{organization size}, but most of the data was
collected in open ended fields. The reason for starting a transformation was
captured in fileds \textit{initial state of organization} and \textit{reasons
for change}. The description of the transformation process was captured in one
field. Recommendations for transformation processes were captured in fields
\textit{reported good practices}, \textit{reported challenges}, and
\textit{lessons learned}. Additional fields that were used but not sufficiently
answered by the primary studies were \textit{satisfaction after transformation},
\textit{effect on onrganization} and \textit{reported measurements before/after
transformation}.

@@

\begin{figure}[h]
  \begin{center}
    \includegraphics[width=1\textwidth]{publicationschart.pdf}
    \caption{Time of publications and transformations}
    \label{fig:publications}
  \end{center}
\end{figure}

@@

\begin{table}[h]
    \begin{tabular}{ l@{ \hskip 0.4cm } l }
        \toprule
        Agile method    & Reported in number of studies   \\ \midrule
        Scrum           & 18 \\ 
        XP              & 7 \\
        Lean            & 5 \\
        Tailoring and combining methods & 9  (indirect mentionings not counted) \\
        \bottomrule
    \end{tabular}
    \caption{Reported use of agile methods}
    \label{table:methods}
\end{table}

The reasons to initiate a change fell into three categories, as summarized in
Table \ref{table:motivations}. The primary reasons for change was a strategical
need to improve the performance of the organization. Especially shortening the
time to market was seen as XXX. In many cases the change was started when
management acknowledged problems in the current way of working.

\begin{table}[h]
    \begin{tabular}{ l@{ \hskip 0.4cm } l }
        \toprule
        Reason                                    & In studies   \\ \midrule
        Need to improve performance               & S8, S11, S15, S16, S17, S24, S27 \\ 
        Need to improve time to market            & S6, S9, S13, S19, S24 \\
        Need to eliminate known process problems  & S5, S6, S10, S13, S19, S21 \\
        \bottomrule
    \end{tabular}
    \caption{Reported resasons for initiating a transformation}
    \label{table:motivations}
\end{table}

@@

\begin{supertabular}{ l p{11.3cm} }
    {[}S1{]} & \bibentry{S1} \\ \shrinkheight{-1.9cm}
    {[}S2{]} & \bibentry{S2} \\ 
    {[}S3{]} & \bibentry{S3} \\ 
    {[}S4{]} & \bibentry{S4} \\ 
    {[}S5{]} & \bibentry{S5} \\ 
    {[}S6{]} & \bibentry{S6} \\ 
    {[}S7{]} & \bibentry{S7} \\ 
    {[}S8{]} & \bibentry{S8} \\ 
    {[}S9{]} & \bibentry{S9} \\ 
    {[}S10{]} & \bibentry{S10} \\ 
    {[}S11{]} & \bibentry{S11} \\ 
    {[}S12{]} & \bibentry{S12} \\ 
    {[}S13{]} & \bibentry{S13} \\ 
    {[}S14{]} & \bibentry{S14} \\ 
    {[}S15{]} & \bibentry{S15} \\ 
    {[}S16{]} & \bibentry{S16} \\ 
    {[}S17{]} & \bibentry{S17} \\ 
    {[}S18{]} & \bibentry{S18} \\ 
    {[}S19{]} & \bibentry{S19} \\ 
    {[}S20{]} & \bibentry{S20} \\ 
    {[}S21{]} & \bibentry{S21} \\ 
    {[}S22{]} & \bibentry{S22} \\ 
    {[}S23{]} & \bibentry{S23} \\ 
    {[}S24{]} & \bibentry{S24} \\ 
    {[}S25{]} & \bibentry{S25} \\ 
    {[}S26{]} & \bibentry{S26} \\ 
    {[}S27{]} & \bibentry{S27} \\ 
    {[}S28{]} & \bibentry{S28} \\ 
    {[}S29{]} & \bibentry{S29} \\ 
    {[}S30{]} & \bibentry{S30} \\ 
\end{supertabular}

>>>>

It is typical for large companies to function according to plan-driven software
engineering models. These models strive for consistent quality and performance
by rigorous planning and definition of process. This kind of approach is however
badly suited for software development, as projects typically encounter
situations that are too hard or impossible to foretell \citegen{Schwaber2002}.
The main problems in plan-driven development are the high cost of changes and
late feedback on quality \citegen{Petersen2010}. Long release cycles, cost of
resopnding to change and distance from customers undermine the  competitiveness
of companies. Agile methods are believed to bring remedy to these problems.

@@

Introducing agile methods in large organizations is more difficult than it is in
small organizations \citegen{Dyba2008}. The difficulty is partly related to size
bringing higher organizational inertia which slows down organizational change
\citegen{Livermore2008}. Agile methods are not founded on the use of individual
tools or practices, but rather on a holistic way of thinking.
Adopting agile requires often change of the entire organizational culture
\citegen{Misra2010}.

One significant difference between small and large adoptions is that large
organizations have more dependencies between projects and teams. This increases
the need for formal documentation and thus reduces agility \citegen{Lindvall2004}.
In addition to inter-team coordination development teams are required to
interact with other organizational units, which are often non-agile of their
nature. For instance, HR may demand individuals to have strictly specified roles
in projects \citegen{Boehm2005}, or a change control board may inhibit use of
continuous integration or refactoring \citegen{Lindvall2004}. All units affected
by the agile transformation need to be informed and heard, and the agile process
must be adjusted according to their needs \citegen{Lindvall2004, Cohn2003,
Boehm2005}.

Agile methods will also affect management and business related functions. A key
challenge is that management must move away from life cycle models and towards
iterative and feature centric models \citegen{Nerur2005}. The focus must be
shifted from a large scale scope to shorter term project planning
\citegen{Misra2010}, as agile methods emphasize that planning is only meaningful
for the near future \citegen{Cohn2003}. However, the lack of planning can be a
concern as business and customer relationships often build on long term
roadmapping. Enabling operation with shorter term planning requires educating
stakeholders and reviewing contracting practices \citegen{Boehm2005}.

@@

The literature review was conducted as an application of Kitchenham's
method for systematic literature review (SLR) \citegen{Kitchenham2007}. The method
was applied to be performed by only one researcher, and to serve as a
preliminary work evaluating the feasibility of a more in-depth review.
Effort was limited by selecting only the following subset of SLR steps:
identification of primary sources, data extraction and synthesis. Steps left out
were the development of a systematic review protocol, strict study quality
assessment, and reviews in the study selection process.

@@

\begin{table}
    \begin{tabular}{ l@{ \hskip 0.4cm } l }
        \toprule
        Facet                  & Keywords   \\ \midrule
        Agile methods          & agile, scrum, lean, xp \\ 
        Organizational change  & transformation, transition, change, migration \\
        Large organization     & enterprise, organization, (large \texttt{AND} scale) \\
        \bottomrule
    \end{tabular}
    \caption{Facets and related search terms used}
    \label{table:searchterms}
\end{table}

@@

@@

The selection process of the primary studies is summarized in Figure
\ref{fig:selection_process}. The database search yielded 1603 matches which were
filtered down to 31 selected primary studies. One study was not available, which
reduced the final number of primary studies to 30.
The 30 primary sources were selected by three criteria: study type, size of
organization, and use of agile methods in software development. Firstly, only
case studies and experience reports on agile software development were included
in this study. Papers discussing agile methods or organizational transformation
on as a general topic were left out. Secondly, the size of the organization
described in the included primary studies was to be large enough to show the
challenges of large scale agile development \citegen{Lindvall2004}. Further,
only studies discussing agile software development were included, leaving out
for instance studies on agile management.

\begin{figure}[t]
  \begin{center}
    \includegraphics[width=0.45\textwidth]{researchprocess}
    \caption{Process of selecting primary studies}
    \label{fig:selection_process}
  \end{center}
\end{figure}

@@

Analysis of the primary studies was guided by using a data extraction form with
open ended fileds. The reason for starting a transformation was captured in
fileds \textit{initial state of organization} and \textit{reasons for change}.
The description of the transformation process was captured in one field.
Recommendations for transformation processes were captured in fields
\textit{reported good practices}, \textit{reported challenges}, and
\textit{lessons learned}.
Additional fields that were used but not sufficiently answered by the primary
studies were \textit{satisfaction after transformation} and \textit{reported
measurements before/after transformation}.

@@

\begin{figure}
  \begin{center}
    \includegraphics[width=1\textwidth]{publicationschart.pdf}
    \caption{Time of publications and transformations}
    \label{fig:publications}
  \end{center}
\end{figure}

@@

\begin{table}
    \begin{tabular}{ l@{ \hskip 0.4cm } l }
        \toprule
        Agile method    & Reported in number of studies   \\ \midrule
        Scrum           & 18 \\ 
        XP              & 7 \\
        Lean            & 5 \\
        Tailoring and combining methods & 9  (indirect mentionings not counted) \\
        \bottomrule
    \end{tabular}
    \caption{Reported use of agile methods}
    \label{table:methods}
\end{table}

The reasons to initiate a change fell into three categories, as summarized in
Table \ref{table:motivations}. The primary reasons for change was a strategical
need to improve the performance of the organization. Especially shortening the
time to market was desired. In many cases the change was started when management
acknowledged problems in the current way of working.

\begin{table}
    \begin{tabular}{ l@{ \hskip 0.4cm } l }
        \toprule
        Reason                                    & In studies   \\ \midrule
        Need to improve performance               & S8, S11, S15, S16, S17, S24, S27 \\ 
        Need to improve time to market            & S6, S9, S13, S19, S24 \\
        Need to eliminate known process problems  & S5, S6, S10, S13, S19, S21 \\
        \bottomrule
    \end{tabular}
    \caption{Reported resasons for initiating a transformation}
    \label{table:motivations}
\end{table}

@@

\begin{supertabular}{ l p{11.3cm} }
    {[}S1{]} & \bibentry{S1} \\ \shrinkheight{-2.1cm}
    {[}S2{]} & \bibentry{S2} \\ 
    {[}S3{]} & \bibentry{S3} \\ 
    {[}S4{]} & \bibentry{S4} \\ 
    {[}S5{]} & \bibentry{S5} \\ 
    {[}S6{]} & \bibentry{S6} \\ 
    {[}S7{]} & \bibentry{S7} \\ 
    {[}S8{]} & \bibentry{S8} \\ 
    {[}S9{]} & \bibentry{S9} \\ 
    {[}S10{]} & \bibentry{S10} \\ 
    {[}S11{]} & \bibentry{S11} \\ 
    {[}S12{]} & \bibentry{S12} \\ 
    {[}S13{]} & \bibentry{S13} \\ 
    {[}S14{]} & \bibentry{S14} \\ 
    {[}S15{]} & \bibentry{S15} \\ 
    {[}S16{]} & \bibentry{S16} \\ 
    {[}S17{]} & \bibentry{S17} \\ 
    {[}S18{]} & \bibentry{S18} \\ 
    {[}S19{]} & \bibentry{S19} \\ 
    {[}S20{]} & \bibentry{S20} \\ 
    {[}S21{]} & \bibentry{S21} \\ 
    {[}S22{]} & \bibentry{S22} \\ 
    {[}S23{]} & \bibentry{S23} \\ 
    {[}S24{]} & \bibentry{S24} \\ 
    {[}S25{]} & \bibentry{S25} \\ 
    {[}S26{]} & \bibentry{S26} \\ 
    {[}S27{]} & \bibentry{S27} \\ 
    {[}S28{]} & \bibentry{S28} \\ 
    {[}S29{]} & \bibentry{S29} \\ 
    {[}S30{]} & \bibentry{S30} \\ 
\end{supertabular}
